Visualizing demographic evolution using geographically inconsistent census data

### 2018-08-18 paper/letter/letter.tex
g2g

\begin{document}
\maketitle
A previous version of this paper was submitted to IEEE VAST 2018. While the full
text of the reviews is attached, we would like to overview the changes made,
leveraging the summary review:

@@

    \reply{We did not introduce a new clustering method. The original paper
    contains such evaluations, we added that explicitly on the text. All the
    communication with the experts is included, verbatim, in the supplementary
    material, with the exception of Patrick Adler, which is a member of the
    Urban Genome Project, but did not collaborate directly in this work.}}

@@

    \reply{We added/expanded discussion about the choices, including the
     subsection about color.}}

    \item{Motivation: the authors fail to clearly demonstrate the motivation of their
    work, probably due to their writing style. (R2, R4)}
    \reply{The introduction was re-written to make this clearer.}

@@

    \reply{The specific region used for the analysis is irrelevant to the
    method, we could have used mixed regions for instance. We adopted CTs
    because it is the most finely grained and volatile data available including
    more than population count. And all the communications with the experts was
    included in the supplementary material.}

@@

    \reply{This change wasn't made because we do not introduce the clustering
    algorithm. For the purposes of this work, any hierarchical clustering would
    work.}}
\end{itemize}

@@

    \reply{The introduction was re-written to make this explicit.}

@@

    \reply{We did not propose a new clusering method. Any hierarchical
    clustering would work, with potentially minimal differences in the result.
    This specific suggestion corresponds to the ClustGeo, an R package that does
    Ward clustering with two distance matrices to express geographic distance.
    That work is mentioned and discussed.}}

@@

    \reply{Colour assignment was indeed challenge for this work, mostly because
    it needs to be used, but there isn't a definitive way to actually do it.
    That was also the reason for artificially limiting the maxium number of
    clusters to eight. There is indeed overlap between the colours used for
    relevance and the clusters, mostly because the clusters used virtually all
    of the colour space. However, we believe that their different meanings are
    clear, since none of the experts, nor the other reviewers, mentioned any
    confusion in this specific aspect. It is not ideal, but we believe it is a
    acceptable approximation. Not ideal, but reasonable enough.}}

    \item{The details of the domain expert interview is missing in the paper.  
    I really appreciate that the authors attached the detailed expert review comments
    and the online demo is available. However, in the main paper, there is a very
    little description of how the expert interviews were conducted. Since there are
    still plenty of space, I would suggest the authors to add more details in Section
    6.}
    \end{enumerate}
    \textbf{Minor issue}:
    The demo video can be improved by providing a more comprehensive walkthrough of a
    case study instead of fragmented videos to explain the functionalities.
}
\end{itemize}

@@

    This paper introduces a combined graph clustering and geospatial visualization
    technique to illustrate the city census data that is potentially inconsistent over
    the space dimension. The technique is generally OK, though not entirely new, and
    can be said as appropriate for the problem attacked. The major strength is the
    detailed case studies on two north American cities and it is appreciated that the
    authors attach a full record of their expert study. On the other hand, the paper
    in its current shape suffers significantly from presentation issues, i.e., beyond
    minor grammar errors, most technical content would benefit from a thorough re-
    written. There is no explicit contribution statement. It seems the contribution on
    visualization is rather limited. The authors are also encouraged to re-submit to a
    GIS venue.

@@

    \reply{Unfortunately the reviewer doesn't mention which is the "most related
    work". I did review some works from Andrienko's group, after all they are
    fairly known for this, but some were not included because they are not
    relevant to the current problem. \\The introduction and related work were
    re-written and significantly expanded. I included more text expliciting
    exactly what was accomplished in this work, which hopefully will make that
    distinction clearer. \\ As the others reviewers mentioned, the tool was
    available not only for the experts, but for the reviewers as well, on an
    anonymous site.}

@@

    \reply{We considered a GIS venue for this work, but the technical part is
    fairly advanced to a general public, requiring significant tangents to
    explain all the involved concepts. Further publications on applied geography
    and demographics venues are planned in the scope of the Urban Genome
    Project, which will cite this publication for the technical details.}}

>>>>

\begin{document}
\maketitle
A previous version of this paper was submitted to IEEE VAST 2018. Before going
to the reviews, I am personally humbly requesting from this new wave of
reviewers: \textbf{please be specific}. I can't really do much with \emph{"most
technical content would benefit from a thorough re-written"} or \emph{"the
writing style and paper organization is hardly a qualified research paper on
visualization"}, especially because I'm following the same pattern of my last
three publications at VAST and TVCG.

The full text of the reviews is attached. We start with the overview of the
changes using the summary review:

@@

    \reply{We did not introduce a new clustering method. The original paper
    contains such evaluations, we added that explicitly on the text. More
    details about the information available to the experts at the time of their
    review was added to the respective section, the supplementary material
    contains the full transcript of our communications.}}

@@

    \reply{We added/expanded the discussion about the visual and
     representational choices, including a five paragraph subsection about
     color.}}

    \item{Presentation: no technical contribution clearly stated, writing style not
    qualified as a good research paper, paper organization can be improved (R2, R4)

    \reply{The paper was re-organized to better reflect our contributions and
     motivation, including a full rewrite of the intro and related work.}}

@@

    \reply{The text was modified to make clear that the specific region used for
    the analysis is irrelevant to the method. We adopted census data because it
    is well explored, and we adopted CTs because it is the most finely grained
    and volatile data available in the censuses that include more than
    population count.}

@@

    \reply{This change wasn't made because we do not introduce the clustering
    algorithm. For the purposes of this work, any hierarchical clustering would
    work. The text was modified to make this point clearer.}}
\end{itemize}

@@

    \reply{The introduction was re-written to make the motivation and our contributions explicit.}

@@

    \reply{I modified the text to make clear that the clustering method is not
    new. Indeed, any hierarchical clustering would work. This specific
    suggestion corresponds to the ClustGeo, an R package that does Ward
    clustering with two distance matrices to express geographic distance. That
    work is mentioned and discussed.}}

@@

    \reply{The colour discussion was expanded. However, we believe the colour
    confusion mentioned is not particular problematic since none of the experts,
    nor the other reviewers, mentioned it. Not ideal, but reasonable enough.}}

    \item{The details of the domain expert interview is missing in the paper.  
    I really appreciate that the authors attached the detailed expert review comments
    and the online demo is available. However, in the main paper, there is a very
    little description of how the expert interviews were conducted. Since there are
    still plenty of space, I would suggest the authors to add more details in Section
    6.
    \reply{Done.}}
    \end{enumerate}
    \textbf{Minor issue}:
    The demo video can be improved by providing a more comprehensive walkthrough of a
    case study instead of fragmented videos to explain the functionalities.

    \reply{These videos are part of the documentation, illustrating specific
    activities.}}
\end{itemize}

@@

    This paper introduces a combined graph clustering and geospatial visualization
    technique to illustrate the city census data that is potentially inconsistent over
    the space dimension. The technique is generally OK, though not entirely new, and
    can be said as appropriate for the problem attacked. The major strength is the
    detailed case studies on two north American cities and it is appreciated that the
    authors attach a full record of their expert study. On the other hand, the paper
    in its current shape suffers significantly from presentation issues, i.e., beyond
    minor grammar errors, most technical content would benefit from a thorough re-
    written. There is no explicit contribution statement. It seems the contribution on
    visualization is rather limited. The authors are also encouraged to re-submit to a
    GIS venue.}

@@

    \reply{Unfortunately the reviewer doesn't mention which is the "most related
    work". I did review some works from Andrienko's group, after all they are
    fairly known for this, but some were not included because they are not
    relevant to the current problem. \\The introduction and related work were
    re-written and significantly expanded. I included more text explaining
    exactly what was accomplished in this work, to make that distinction
    clearer. \\ As the others reviewers mentioned, quite obviously the tool was
    available to the experts, and for the reviewers as well, on an anonymous
    site.}

@@

    \reply{We considered a GIS venue for this work, but the technical part is
    fairly advanced to a general public, requiring significant tangents to
    explain all the involved concepts. Further publications on applied geography
    and demographics venues are planned in the scope of the Urban Genome
    Project, which will cite this publication for the technical details.}}
\end{itemize}

### 2018-08-18 paper/study.tex
g2g

\section{Expert feedback}
%This could be a big step toward better understanding of evolving urban morphology.
To assess the proposed method, we contacted academic and industry experts in
sociology and urban sciences. We focused our inquires on the utility of the
prototype, and if they found anything interesting on the data using it. The
supplementary material contains the communication with five experts, verbatim,
identified from A to E.

>>>>

\section{Expert feedback}
We contacted academic and industry experts in sociology and urban sciences for
their appreciation of our methodology. They had access to the prototype tool, a
descriptive documentation of the features (included in the supplementary
material), and a sequence of documentation videos illustrating how to perform
specific tasks. The documentation explains which datasets are used, how the data
is represented and processed, including that there is no geographic
harmonization. We focused our inquires on the results obtained, asking if they
found anything interesting on the data. The message sent and their full response
is included in the supplementary material. Each of the five experts is
identified by a letter, from A to E.
